Centers and centralizers of higher enveloping algebras of Lie algebras

### 2020-09-09 draft.tex
slight rearrangement

Recently, a generalization of the concept of the enveloping algebra of a Lie algebra has emerged: 
each dg Lie algebra $\fg$ has an enveloping $\EE_n$ algebra $\UU_n\fg$, for every $n$, 
so that $\UU_1\fg$ is the usual enveloping algebra $U\fg$. 
Building on prior work \cite{BD, AF} [Owen's thesis?], 
Knudsen \cite{Knudsen} developed these ideas extensively and offered a useful factorization algebra model for the enveloping $\EE_n$ algebra. 
Our central goal here is to generalize the some well-known results about the enveloping algebra $U\fg$ to $\UU_n\fg$, for all $n$, such as
\begin{enumerate}
\item[(1)] the (underived) center of $U\sl_{N+1}$ is isomorphic to $(\Sym(\CC^{N}))^{S_N}$, where $\CC^{N}$ denotes the Cartan subalgebra of $\sl_{N+1}$ and $S_N$ is the Weyl group of $SL_{N+1}$, and
\item[(2)] the derived center of the enveloping algebra $U\fg$ is given by derived invariants of the adjoint action of $\fg$ on $U\fg$. That is to say $\Hoch^*(U\fg,U\fg) \simeq \clies(\fg, U\fg)$.
\end{enumerate}
\owen{And keep going}
%
%Our results pertains to (1) by taking $N \to \infty$, and to (2) through the derived {\em centralizer} of the map $\UU_n\sl_\infty \hookrightarrow \UU_n\gl_\infty(A)$, where $A$ is a dg algebra. Note that the usual derived center is the centralizer of the identity map $\UU_n\fg \to \UU_n\fg$.
%
%We will review these notions and describe our results in detail Min the next section. When $n=1$,  \owen{state the classical thing here!}
%
%Our primary tool here is the Loday-Quillen-Tsygan (LQT) theorem which describes the large $N$ limit of the Lie algebra homology of $\gl_N(A)$, for a dg associative algebra $A$. Since the enveloping $\EE_n$ algebra of a Lie algebra can be constructed as a locally constant factorization algebra made out of the shifted Chevalley-Eilenberg chains \cite{}, the relevance of the LQT theorem in this context  should be apparent. From this point of view, one can think of our main result as an $\EE_n$ generalization of the LQT theorem. We hope these ideas have exciting applications in quantum field theory and random matrix theory, as well as to other large $N$ limit phenomena.

@@

\owen{Some easy but important situations to examine:}
\begin{itemize}
\item When $n=0$, we see naive endomorphisms arise.
\item When $n=1$, we recover standard Koszul duality and Hochschild cohomology examples.
\item When $n=1$, unpack in detail how the convolution picture matches the classic view of Koszul duality for associative (co)algebras.
\item Consider $\fg$ {\em abelian} for arbitrary $n$ and compare to the $\EE_n$ HKR theorem.
\item Compare with $\EE_n$ HKR theorem for $\clies(\fg)$ via $\EE_n$ Koszul duality.
\item Use these facts to compute factorization homology of centralizers on manifolds with boundary --- kind of like Chern-Simons states!
\end{itemize}

Having an explicit presentation for the centralizer of a map of $\EE_n$ algebras leads to a calculation of factorization homology.
The enveloping $\EE_n$ algebra is the homotopy $O(n)$-fixed points, or a $\mathrm{Disk}_n$ algebra in the sense of~\cite{AF},
and hence determines a factorization algebra on any un-oriented $n$-dimensional manifold \cite{Knudsen}.
Therefore, for a closed $n$-dimensional manifold $M$, one has
\[
\int_M {\frak Z}_n(\UU_n \rho) \simeq \clies(\fg \otimes \Omega^*(M), \cliels( \fh \otimes \Omega^*(M))).
\]
Since $\Omega^*(S^n) \simeq \CC[\epsilon_n]$, where $\epsilon_n$ is a formal variable with $|\epsilon_n| = n$ and $\epsilon_n^2 = 0$, we have
\[
\int_{S^n} {\frak Z}_n(\UU_n \rho) \simeq \clies(\fg[\epsilon_n], \cliels( \fh[\epsilon_n]))
\]
For $n=1$, this recovers a theorem of Tamarkin and Tsygan \cite{TT} on the Hochschild cohomology of differential operators on a Lie group $G$;
see Example~\ref{eg: TT}.

@@

\owen{I think we can conjecture here the general statement that we almost proved and sketch an approach.}

@@

Our model also gives us a simple description of the factorization homology of the derived center along the circle~$S^1$:
\begin{align*}
\Hoch_*(\Hoch^*(U \fg, U\fg)) 
&\simeq \int_{S^1} {\frak Z}(U(\fg)) \\
&\simeq \clies(\fg[\epsilon], \cliels( \fg[\epsilon])),
\end{align*}
where $\epsilon$ is a formal variable of degree~$1$. 
In the last step we use the fact that the circle is formal as a rational space.
\brian{We should compare with Tamarkin-Tsygan's noncommutative calculus (cf. section 2.6).
We could exhibit Theorem 2.7.1 explicitly in this case for the $\EE_n$ case with enveloping algebras.}

>>>>

Recently, a generalization of the concept of the enveloping algebra of a Lie algebra has emerged: 
each dg Lie algebra $\fg$ has an enveloping $\EE_n$ algebra $\UU_n\fg$, for every $n$, 
so that $\UU_1\fg$ is the usual enveloping algebra $U\fg$. 
Building on prior work \cite{BD, AF} [Owen's thesis?], 
Knudsen \cite{Knudsen} developed these ideas extensively and offered a useful factorization algebra model for the enveloping $\EE_n$ algebra. 
Our central goal here is to generalize the some well-known results about the enveloping algebra $U\fg$ to $\UU_n\fg$, for all $n$, such as
\begin{enumerate}
\item[(1)] the (underived) center of $U\sl_{N+1}$ is isomorphic to $(\Sym(\CC^{N}))^{S_N}$, where $\CC^{N}$ denotes the Cartan subalgebra of $\sl_{N+1}$ and $S_N$ is the Weyl group of $SL_{N+1}$, and
\item[(2)] the derived center of the enveloping algebra $U\fg$ is given by derived invariants of the adjoint action of $\fg$ on $U\fg$. That is to say $\Hoch^*(U\fg,U\fg) \simeq \clies(\fg, U\fg)$.
\end{enumerate}
\owen{And keep going just a little.}
%
%Our results pertains to (1) by taking $N \to \infty$, and to (2) through the derived {\em centralizer} of the map $\UU_n\sl_\infty \hookrightarrow \UU_n\gl_\infty(A)$, where $A$ is a dg algebra. Note that the usual derived center is the centralizer of the identity map $\UU_n\fg \to \UU_n\fg$.
%
%We will review these notions and describe our results in detail Min the next section. When $n=1$,  \owen{state the classical thing here!}
%
%Our primary tool here is the Loday-Quillen-Tsygan (LQT) theorem which describes the large $N$ limit of the Lie algebra homology of $\gl_N(A)$, for a dg associative algebra $A$. Since the enveloping $\EE_n$ algebra of a Lie algebra can be constructed as a locally constant factorization algebra made out of the shifted Chevalley-Eilenberg chains \cite{}, the relevance of the LQT theorem in this context  should be apparent. From this point of view, one can think of our main result as an $\EE_n$ generalization of the LQT theorem. We hope these ideas have exciting applications in quantum field theory and random matrix theory, as well as to other large $N$ limit phenomena.

@@

\owen{We could describe more that we do.}

@@

\owen{If we want, we could conjecture here the general statement that we almost proved (for $f: \UU_n \fg \to B$, $\frak{Z}_n(f) \simeq \clies(\fg, B_f)$) and maybe even sketch an approach.}

@@

Our model also gives us a simple description of the factorization homology of the derived center along the circle~$S^1$:
\begin{align*}
\Hoch_*(\Hoch^*(U \fg, U\fg)) 
&\simeq \int_{S^1} {\frak Z}(U(\fg)) \\
&\simeq \clies(\fg[\epsilon], \cliels( \fg[\epsilon])),
\end{align*}
where $\epsilon$ is a formal variable of degree~$1$. 
In the last step we use the fact that the circle is formal as a rational space.
\brian{For $n=1$, this recovers a theorem of Tamarkin and Tsygan \cite{TT} on the Hochschild cohomology of differential operators on a Lie group $G$;
see Example~\ref{eg: TT}(cf. section 2.6).
We could describe an $\EE_n$ generalization?}
